---
title: "Data overview for Scorecards project"
author: "Zuzanna Kostecka, Karolina Solarska"
date: "2023-06-09"
output: html_document
---
```{r}
library(knitr)
# install.packages("readr")
library(readr)
library(corrplot)
```

```{r}
# Karolina:
# setwd("C:/Users/karla/OneDrive/Pulpit/Credit Risk/Projekt")
# app_train3<-read_csv("app_test3.csv")
# app_test3<-read_csv("app_train3.csv")

# Zuzia:
setwd("/Users/zuzanna/Desktop/Studia/CreditRisk/projekt")
load("dane/app_train_prep.RData")
load("dane/app_test_prep.RData")

```

# Introduction

This is an overview of our initially prepared data. We extracted the applicant's 
profiles from [Kaggle](https://www.kaggle.com/competitions/home-credit-default-risk/data).

The summary of Train Data.
```{r, echo=TRUE}
summary(app_train3)
```
<br>
<br>

The summary of Test Data.
```{r, echo=TRUE}
summary(app_test3)
```
<br>
<br>

# 0. TARGET

```{r, echo=TRUE}
unique(app_train3$TARGET)
```

```{r}
app_train3$TARGET <- as.numeric(app_train3$TARGET)
summary(app_train3$TARGET)
targetTR <- table(app_train3$TARGET)
barplot(targetTR, main = "Target variable is train sample",
        ylab = "Count",
        names.arg = c("0", "1"),
        col = c("grey", "orange"))
```

# 1. NAME_CONTRACT_TYPE

We can observe the contract type is a text variable. We need to transform it to
categorical data type. 
<br>
```{r, echo=TRUE}
unique(app_train3$NAME_CONTRACT_TYPE)
unique(app_test3$NAME_CONTRACT_TYPE)
```
<br>

After transformation: 

Cash loans = 0 and Revolving loans = 1
```{r, echo=FALSE}
app_train3$NAME_CONTRACT_TYPE <- ifelse(app_train3$NAME_CONTRACT_TYPE == "Cash loans", 0,
                  ifelse(app_train3$NAME_CONTRACT_TYPE == "Revolving loans", 1,
                  sample(c(0, 1), length(app_train3$NAME_CONTRACT_TYPE), replace = TRUE)))
app_test3$NAME_CONTRACT_TYPE <- ifelse(app_test3$NAME_CONTRACT_TYPE == "Cash loans", 0,
                  ifelse(app_test3$NAME_CONTRACT_TYPE == "Revolving loans", 1,
                  sample(c(0, 1), length(app_test3$NAME_CONTRACT_TYPE), replace = TRUE)))

# Assign numeric values to the factor levels based on the mapping
app_train3$NAME_CONTRACT_TYPE <- as.integer(app_train3$NAME_CONTRACT_TYPE)
app_test3$NAME_CONTRACT_TYPE <- as.integer(app_test3$NAME_CONTRACT_TYPE)

# Now we can verify the changes
unique(app_test3$NAME_CONTRACT_TYPE)

is.numeric(app_train3$NAME_CONTRACT_TYPE)
is.numeric(app_test3$NAME_CONTRACT_TYPE)

loanTR <- table(app_train3$NAME_CONTRACT_TYPE)
loanTR
barplot(loanTR, main = "Counts of Loan Types in train sample",
        ylab = "Count",
        names.arg = c("Cash Loan", "Revolving Loan"),
        col = c("grey", "orange"))

loanTE <- table(app_test3$NAME_CONTRACT_TYPE)
loanTE
barplot(loanTE, main = "Counts of Loan Types in test sample",
        ylab = "Count",
        names.arg = c("Cash Loan", "Revolving Loan"),
        col = c("grey", "orange"))


```

<br>
<br>
<br>

# 2. CODE_GENDER

We can observe the gender is a text variable. We need to transform it to
binary data type. 
<br>
```{r, echo=TRUE}
unique(app_train3$CODE_GENDER)
unique(app_test3$CODE_GENDER)
```
<br>

After transformation: 

```{r, echo=FALSE}
app_train3$CODE_GENDER <- ifelse(app_train3$CODE_GENDER == "M", 0,
                                       ifelse(app_train3$CODE_GENDER == "F", 1,
                                              sample(c(0, 1), length(app_train3$CODE_GENDER), replace = TRUE)))

unique(app_train3$CODE_GENDER)

app_test3$CODE_GENDER <- ifelse(app_test3$CODE_GENDER == "M", 0, 1)
unique(app_test3$CODE_GENDER)

is.numeric(app_train3$CODE_GENDER)
is.numeric(app_test3$CODE_GENDER)

genderTR <- table(app_train3$CODE_GENDER)

barplot(genderTR, main = "Gender participation in train sample",
        ylab = "Count",
        names.arg = c("Male", "Female"),
        col = c("grey", "orange"))

genderTE <- table(app_test3$CODE_GENDER)

barplot(genderTE, main = "Gender participation in test sample",
        ylab = "Count",
        names.arg = c("Male", "Female"),
        col = c("grey", "orange"))
```

<br>
<br>
<br>

# 3. FLAG_OWN_CAR

We can observe the owning car is a text variable. We need to transform it to
binary data type. 
<br>
```{r, echo=TRUE}
unique(app_train3$FLAG_OWN_CAR)
unique(app_test3$FLAG_OWN_CAR)
```
<br>

After transformation: 

```{r, echo=FALSE}
app_train3$FLAG_OWN_CAR <- ifelse(app_train3$FLAG_OWN_CAR == "N", 0, 1)
unique(app_train3$FLAG_OWN_CAR)

app_test3$FLAG_OWN_CAR <- ifelse(app_test3$FLAG_OWN_CAR == "N", 0, 1)
unique(app_test3$FLAG_OWN_CAR)

is.numeric(app_train3$FLAG_OWN_CAR)
is.numeric(app_test3$FLAG_OWN_CAR)

carTR <- table(app_train3$FLAG_OWN_CAR)

barplot(carTR, main = "Car owners in train sample",
        ylab = "Count",
        names.arg = c("Do not own a car", "Own a car"),
        col = c("grey", "orange"))

carTE <- table(app_test3$FLAG_OWN_CAR)

barplot(carTE, main = "Car owners in test sample",
        ylab = "Count",
        names.arg = c("Do not own a car", "Own a car"),
        col = c("grey", "orange"))

```

<br>
<br>
<br>

# 4. FLAG_OWN_REALTY

We can observe the owning a flat or house is a text variable. We need to transform it to
binary data type. 
<br>
```{r, echo=TRUE}
unique(app_train3$FLAG_OWN_REALTY)
unique(app_test3$FLAG_OWN_REALTY)
```
<br>

After transformation: 

```{r, echo=FALSE}
app_train3$FLAG_OWN_REALTY <- ifelse(app_train3$FLAG_OWN_REALTY == "N", 0, 1)
unique(app_train3$FLAG_OWN_REALTY)

app_test3$FLAG_OWN_REALTY <- ifelse(app_test3$FLAG_OWN_REALTY == "N", 0, 1)
unique(app_test3$FLAG_OWN_REALTY)

is.numeric(app_train3$FLAG_OWN_REALTY)
is.numeric(app_test3$FLAG_OWN_REALTY)

realtyTR <- table(app_train3$FLAG_OWN_REALTY)

barplot(realtyTR, main = "Flat/house owners in train sample",
        ylab = "Count",
        names.arg = c("Do not own a flat/house", "Own a flat/house"),
        col = c("grey", "orange"))

realtyTE <- table(app_test3$FLAG_OWN_CAR)

barplot(realtyTE, main = "Flat/house owners in test sample",
        ylab = "Count",
        names.arg = c("Do not own a flat/house", "Own a flat/house"),
        col = c("grey", "orange"))

```

<br>
<br>
<br>

# 5. CNT_CHILDREN


<br>
```{r, echo=TRUE}
unique(app_train3$CNT_CHILDREN)
unique(app_test3$CNT_CHILDREN)
```
<br>

```{r, echo=FALSE}
childTR <- table(app_train3$CNT_CHILDREN)

barplot(childTR, main = "Number of children in train sample",
        ylab = "Count"
        )

childTE <- table(app_test3$CNT_CHILDREN)

barplot(childTE, main = "Number of children  in test sample",
        ylab = "Count"
        )
```

<br> 
<br>

In this case, we will treat 4 children and more as multiple families. We need to 
cap the data.

```{r, echo = FALSE}
# Replace values above 4 with 4 in CNT_CHILDREN column
app_train3$CNT_CHILDREN <- ifelse(app_train3$CNT_CHILDREN > 4, 4, app_train3$CNT_CHILDREN)
app_test3$CNT_CHILDREN <- ifelse(app_test3$CNT_CHILDREN > 4, 4, app_test3$CNT_CHILDREN)

childTR <- table(app_train3$CNT_CHILDREN)

barplot(childTR, main = "Number of children in train sample after capping",
        ylab = "Count"
        )

childTE <- table(app_test3$CNT_CHILDREN)

barplot(childTE, main = "Number of children  in test sample after capping",
        ylab = "Count"
        )
```

Transform the variable to binary.

```{r, echo = FALSE}
# Replace values above 0 with 1 in CNT_CHILDREN column
app_train3$CNT_CHILDREN <- ifelse(app_train3$CNT_CHILDREN > 0, 1, app_train3$CNT_CHILDREN)
app_test3$CNT_CHILDREN <- ifelse(app_test3$CNT_CHILDREN > 0, 1, app_test3$CNT_CHILDREN)

childTR <- table(app_train3$CNT_CHILDREN)

barplot(log(childTR), main = "Number of children in train sample after capping to 0-1",
        ylab = "Count"
        )

childTE <- table(app_test3$CNT_CHILDREN)

barplot(childTE, main = "Number of children  in test sample after capping to 0-1",
        ylab = "Count"
        )
```

On this stage, our children variable is balanced for train sample and imbalanced
for test sample.

<br>
<br>
<br>


# 6. AMT_INCOME_TOTAL

```{r, echo=FALSE}
summary(app_test3$AMT_INCOME_TOTAL)
summary(app_train3$AMT_INCOME_TOTAL)

# Plot the distribution of AMT_INCOME_TOTAL
hist(app_test3$AMT_INCOME_TOTAL, 
     main = "Distribution of Income in test sample",
     xlab = "Income Total")

hist(app_train3$AMT_INCOME_TOTAL, 
     main = "Distribution of income in train sample",
     xlab = "Income Total")
```
<br>
<br>
We can observe that the outliers are only form the right side. We use winsorization.
```{r, echo=FALSE}
# Winsorize AMT_INCOME_TOTAL from the right side by 5%

# Calculate the 95th percentile
p95 <- quantile(app_test3$AMT_INCOME_TOTAL, 0.95, na.rm = TRUE)

# Winsorize AMT_INCOME_TOTAL from the right side by 5%
app_test3$AMT_INCOME_TOTAL[app_test3$AMT_INCOME_TOTAL > p95] <- p95
app_train3$AMT_INCOME_TOTAL[app_train3$AMT_INCOME_TOTAL > p95] <- p95


hist(app_test3$AMT_INCOME_TOTAL, 
     main = "Distribution of Income in test sample after winsorization",
     xlab = "Income Total")

hist(app_train3$AMT_INCOME_TOTAL, 
     main = "Distribution of income in train sample after winsorization",
     xlab = "Income Total")
```

# 7. AMT_CREDIT

```{r, echo=FALSE}
summary(app_test3$AMT_CREDIT)
summary(app_train3$AMT_CREDIT)

# Plot the distribution of AMT_CREDIT
hist(app_test3$AMT_CREDIT, 
     main = "Distribution of Credit Application in test sample",
     xlab = "Income Total")

hist(app_train3$AMT_CREDIT, 
     main = "Distribution of Credit Application in train sample",
     xlab = "Income Total")
```
<br>
<br>

Applying logarithmic transformation.

```{r}

app_train3$AMT_CREDIT <- log(app_train3$AMT_CREDIT)
app_test3$AMT_CREDIT <- log(app_test3$AMT_CREDIT)

hist(app_test3$AMT_CREDIT, 
     main = "Distribution of CA in test sample after logarithimic transformation",
     xlab = "Credit Total")

hist(app_train3$AMT_CREDIT, 
     main = "Distribution of CA in train sample after logarithimic transformation",
     xlab = "Credit Total")
```

# 8. AMT_ANNUITY

```{r, echo=FALSE}
summary(app_test3$AMT_ANNUITY)
summary(app_train3$AMT_ANNUITY)

# Plot the distribution of AMT_ANNUITY
hist(app_test3$AMT_ANNUITY, 
     main = "Distribution of annuity in test sample",
     xlab = "Annuity Total")

hist(app_train3$AMT_ANNUITY, 
     main = "Distribution of annuity in train sample",
     xlab = "Annuity Total")
```

<br>

We need to handle NA's (by imputation), and distribution applying logarithmic transformation.
```{r, echo=FALSE}

# Create a subset of the data without missing values
no_missing <- app_test3$AMT_ANNUITY[!is.na(app_test3$AMT_ANNUITY)]

# Generate random values from the subset distribution
random_values <- sample(no_missing, sum(is.na(app_test3$AMT_ANNUITY)), replace = TRUE)

# Replace missing values with random values
app_test3$AMT_ANNUITY[is.na(app_test3$AMT_ANNUITY)] <- random_values

app_test3$AMT_ANNUITY <- log(app_test3$AMT_ANNUITY)


# Create a subset of the data without missing values
no_missing <- app_train3$AMT_ANNUITY[!is.na(app_train3$AMT_ANNUITY)]

# Generate random values from the subset distribution
random_values <- sample(no_missing, sum(is.na(app_train3$AMT_ANNUITY)), replace = TRUE)

# Replace missing values with random values
app_train3$AMT_ANNUITY[is.na(app_train3$AMT_ANNUITY)] <- random_values

app_train3$AMT_ANNUITY <- log(app_train3$AMT_ANNUITY)


# Plot the distribution of AMT_ANNUITY
hist(app_test3$AMT_ANNUITY, 
     main = "Distribution of annuity in test sample after transformation",
     xlab = "Annuity Total")

hist(app_train3$AMT_ANNUITY, 
     main = "Distribution of annuity in train sample after transformation",
     xlab = "Annuity Total")

```
<br>

# 9. AMT_GOODS_PRICE

```{r}
# Summary statistics of AMT_GOODS_PRICE
summary(app_test3$AMT_GOODS_PRICE)
summary(app_train3$AMT_GOODS_PRICE)

# Create a subset of the data without missing values
no_missing_train <- app_train3$AMT_GOODS_PRICE[!is.na(app_train3$AMT_GOODS_PRICE)]
no_missing_test <- app_test3$AMT_GOODS_PRICE[!is.na(app_test3$AMT_GOODS_PRICE)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$AMT_GOODS_PRICE)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$AMT_GOODS_PRICE)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$AMT_GOODS_PRICE[is.na(app_train3$AMT_GOODS_PRICE)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$AMT_GOODS_PRICE[is.na(app_test3$AMT_GOODS_PRICE)] <- random_values_test

# Plot the distribution of AMT_GOODS_PRICE
hist(app_test3$AMT_GOODS_PRICE,
     main = "Distribution of Goods Price in Test Sample",
     xlab = "Goods Price")

hist(app_train3$AMT_GOODS_PRICE,
     main = "Distribution of Goods Price in Train Sample",
     xlab = "Goods Price")

# Boxplots
boxplot(app_test3$AMT_GOODS_PRICE,
        main = "Boxplot of Goods Price in Test Sample",
        ylab = "Goods Price")

boxplot(app_train3$AMT_GOODS_PRICE,
        main = "Boxplot of Goods Price in Train Sample",
        ylab = "Goods Price")

```
<br>

Apply the logarithmic transformation.
<br>

```{r}
app_test3$AMT_GOODS_PRICE <- log(app_test3$AMT_GOODS_PRICE)
app_train3$AMT_GOODS_PRICE <- log(app_train3$AMT_GOODS_PRICE)

# Plot the distribution of AMT_GOODS_PRICE
hist(app_test3$AMT_GOODS_PRICE,
     main = "Distribution of Goods Price in Test Sample after transformation",
     xlab = "Goods Price")

hist(app_train3$AMT_GOODS_PRICE,
     main = "Distribution of Goods Price in Train Sample after transformation",
     xlab = "Goods Price")
```
<br>
<br>

# 10. NAME_TYPE_SUITE

```{r}
# Summary statistics of NAME_TYPE_SUITE
summary(app_test3$NAME_TYPE_SUITE)
summary(app_train3$NAME_TYPE_SUITE)

# We can observe the NAME_TYPE_SUITE is a text variable. We need to transform it to categorical data type.
unique(app_train3$NAME_TYPE_SUITE)
unique(app_test3$NAME_TYPE_SUITE)

# So we will create a mapping of categories to numeric values
category_mapping <- c("Unaccompanied" = 1, "Family" = 2, "Spouse, partner" = 3, "Group of people" = 4, "Other_B" = 5, "Children" = 6, "Other_A" = 7)

# Then convert the "NAME_TYPE_SUITE" column to a factor with the specified categories
app_train3$NAME_TYPE_SUITE <- factor(app_train3$NAME_TYPE_SUITE, levels = names(category_mapping))
app_test3$NAME_TYPE_SUITE <- factor(app_test3$NAME_TYPE_SUITE, levels = names(category_mapping))

# Assign numeric values to the factor levels based on the mapping
app_train3$NAME_TYPE_SUITE <- as.integer(app_train3$NAME_TYPE_SUITE)
app_test3$NAME_TYPE_SUITE <- as.integer(app_test3$NAME_TYPE_SUITE)

# Now we can verify the changes
unique(app_test3$NAME_TYPE_SUITE)

# and we saw hat we need to handle missing data
# Create a subset of the data without missing values
no_missing_train <- app_train3$NAME_TYPE_SUITE[!is.na(app_train3$NAME_TYPE_SUITE)]
no_missing_test <- app_test3$NAME_TYPE_SUITE[!is.na(app_test3$NAME_TYPE_SUITE)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$NAME_TYPE_SUITE)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$NAME_TYPE_SUITE)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$NAME_TYPE_SUITE[is.na(app_train3$NAME_TYPE_SUITE)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$NAME_TYPE_SUITE[is.na(app_test3$NAME_TYPE_SUITE)] <- random_values_test

# Plot the distribution of AMT_ANNUITY
hist(app_test3$NAME_TYPE_SUITE, 
     main = "Distribution of type suite in test sample",
     xlab = "Type suite")

hist(app_train3$NAME_TYPE_SUITE, 
     main = "Distribution of type suite in train sample",
     xlab = "Type suite")


```

<br>

After checking the distribution, we decide to parse the NAME_TYPE_SUITE as binary variable.


```{r}
app_test3$NAME_TYPE_SUITE <- ifelse(app_test3$NAME_TYPE_SUITE == '1', 0, 1)
app_train3$NAME_TYPE_SUITE <- ifelse(app_train3$NAME_TYPE_SUITE == '1', 0, 1)

hist(app_test3$NAME_TYPE_SUITE, 
     main = "Distribution of type suite in test sample",
     xlab = "Type suite")

hist(app_train3$NAME_TYPE_SUITE, 
     main = "Distribution of type suite in train sample",
     xlab = "Type suite")

```

<br>

# 11. NAME_EDUCATION_TYPE

```{r}
# Summary statistics of NAME_EDUCATION_TYPE
summary(app_test3$NAME_EDUCATION_TYPE)
summary(app_train3$NAME_EDUCATION_TYPE)

# We can observe the NAME_EDUCATION_TYPE is a text variable. We need to transform it to categorical data type.
unique(app_train3$NAME_EDUCATION_TYPE)
unique(app_test3$NAME_EDUCATION_TYPE)

# So we will create a mapping of categories to numeric values
category_mapping <- c("Lower secondary" = 1, "Secondary / secondary special" = 2, "Incomplete higher" = 3, "Higher education" = 4, "Academic degree" = 5)

# Then convert the "NAME_EDUCATION_TYPE" column to a factor with the specified categories
app_train3$NAME_EDUCATION_TYPE <- factor(app_train3$NAME_EDUCATION_TYPE, levels = names(category_mapping))
app_test3$NAME_EDUCATION_TYPE <- factor(app_test3$NAME_EDUCATION_TYPE, levels = names(category_mapping))

# Assign numeric values to the factor levels based on the mapping
app_train3$NAME_EDUCATION_TYPE <- as.integer(app_train3$NAME_EDUCATION_TYPE)
app_test3$NAME_EDUCATION_TYPE <- as.integer(app_test3$NAME_EDUCATION_TYPE)

# Now we can verify the changes
unique(app_test3$NAME_EDUCATION_TYPE)

# and we saw hat we need to handle missing data
# Create a subset of the data without missing values
no_missing_train <- app_train3$NAME_EDUCATION_TYPE[!is.na(app_train3$NAME_EDUCATION_TYPE)]
no_missing_test <- app_test3$NAME_EDUCATION_TYPE[!is.na(app_test3$NAME_EDUCATION_TYPE)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$NAME_EDUCATION_TYPE)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$NAME_EDUCATION_TYPE)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$NAME_EDUCATION_TYPE[is.na(app_train3$NAME_EDUCATION_TYPE)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$NAME_EDUCATION_TYPE[is.na(app_test3$NAME_EDUCATION_TYPE)] <- random_values_test

# Plot the distribution of AMT_ANNUITY
hist(app_test3$NAME_EDUCATION_TYPE, 
     main = "Distribution of education type in test sample",
     xlab = "education type")

hist(app_train3$NAME_EDUCATION_TYPE, 
     main = "Distribution of education type in train sample",
     xlab = "education type")
```

After checking the distribution, we decide to parse the NAME_EDUCATION_TYPE as binary variable.
Higher education and above is 1 and lower is 0.

```{r}
app_test3$NAME_EDUCATION_TYPE <- ifelse(app_test3$NAME_EDUCATION_TYPE == '1', 0, 
                (ifelse(app_test3$NAME_EDUCATION_TYPE == '2', 0, 1)))
app_train3$NAME_EDUCATION_TYPE <- ifelse(app_train3$NAME_EDUCATION_TYPE == '1', 0, 
                (ifelse(app_train3$NAME_EDUCATION_TYPE == '2', 0, 1)))

hist(app_test3$NAME_EDUCATION_TYPE, 
     main = "Distribution of education in test sample",
     xlab = "Educ")

hist(app_train3$NAME_EDUCATION_TYPE, 
     main = "Distribution of education in train sample",
     xlab = "Educ")

```

<br>
<br>

# 12. NAME_FAMILY_STATUS

```{r}
# Summary statistics of NAME_FAMILY_STATUS
summary(app_test3$NAME_FAMILY_STATUS)
summary(app_train3$NAME_FAMILY_STATUS)

# We can observe the NAME_FAMILY_STATUS is a text variable. We need to transform it to categorical data type.
unique(app_train3$NAME_FAMILY_STATUS)
unique(app_test3$NAME_FAMILY_STATUS)

# So we will create a mapping of categories to numeric values
category_mapping <- c("Single / not married" = 1, "Civil marriage" = 2, "Married" = 3, "Separated" = 4, "Widow" = 5)

# Then convert the "NAME_FAMILY_STATUS" column to a factor with the specified categories
app_train3$NAME_FAMILY_STATUS <- factor(app_train3$NAME_FAMILY_STATUS, levels = names(category_mapping))
app_test3$NAME_FAMILY_STATUS <- factor(app_test3$NAME_FAMILY_STATUS, levels = names(category_mapping))

# Assign numeric values to the factor levels based on the mapping
app_train3$NAME_FAMILY_STATUS <- as.integer(app_train3$NAME_FAMILY_STATUS)
app_test3$NAME_FAMILY_STATUS <- as.integer(app_test3$NAME_FAMILY_STATUS)

# Now we can verify the changes
unique(app_test3$NAME_FAMILY_STATUS)

# and we saw hat we need to handle missing data
# Create a subset of the data without missing values
no_missing_train <- app_train3$NAME_FAMILY_STATUS[!is.na(app_train3$NAME_FAMILY_STATUS)]
no_missing_test <- app_test3$NAME_FAMILY_STATUS[!is.na(app_test3$NAME_FAMILY_STATUS)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$NAME_FAMILY_STATUS)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$NAME_FAMILY_STATUS)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$NAME_FAMILY_STATUS[is.na(app_train3$NAME_FAMILY_STATUS)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$NAME_FAMILY_STATUS[is.na(app_test3$NAME_FAMILY_STATUS)] <- random_values_test

# Plot the distribution of AMT_ANNUITY
hist(app_test3$NAME_FAMILY_STATUS, 
     main = "Distribution of family status in test sample",
     xlab = "family status")

hist(app_train3$NAME_FAMILY_STATUS, 
     main = "Distribution of family status in train sample",
     xlab = "family status")

```

<br>
<br>

# 13. DAYS_EMPLOYED

```{r}
# Summary statistics of DAYS_EMPLOYED
summary(app_test3$DAYS_EMPLOYED)
summary(app_train3$DAYS_EMPLOYED)

sum(is.na(app_test3$DAYS_EMPLOYED))
sum(is.na(app_train3$DAYS_EMPLOYED))

# Plot the distribution of DAYS_EMPLOYED
hist(app_test3$DAYS_EMPLOYED, 
     main = "Distribution of employed days in test sample",
     xlab = "employed days")

hist(app_train3$DAYS_EMPLOYED, 
     main = "Distribution of employed days in train sample",
     xlab = "employed days")
```

<br>
<br>

# 14. REGION_RATING_CLIENT

```{r}
# Summary statistics of REGION_RATING_CLIENT
summary(app_test3$REGION_RATING_CLIENT)
summary(app_train3$REGION_RATING_CLIENT)

sum(is.na(app_test3$REGION_RATING_CLIENT))
sum(is.na(app_train3$REGION_RATING_CLIENT))

# Plot the distribution of REGION_RATING_CLIENT
hist(app_test3$REGION_RATING_CLIENT, 
     main = "Distribution of region rating client in test sample",
     xlab = "region rating client")

hist(app_train3$REGION_RATING_CLIENT, 
     main = "Distribution of region rating client in train sample",
     xlab = "region rating client")
```

<br>
<br>

# 15. DOCUMENT

```{r}
# Summary statistics of DOCUMENT
summary(app_test3$DOCUMENT)
summary(app_train3$DOCUMENT)

sum(is.na(app_test3$DOCUMENT))
sum(is.na(app_train3$DOCUMENT))

app_test3$DOCUMENT <- ifelse(app_test3$DOCUMENT > 0, 1, 0)
app_train3$DOCUMENT <- ifelse(app_train3$DOCUMENT > 0, 1, 0)

# Plot the distribution of DOCUMENT
hist(app_test3$DOCUMENT, 
     main = "Distribution of DOCUMENT in test sample",
     xlab = "DOCUMENT")

hist(app_train3$DOCUMENT, 
     main = "Distribution of DOCUMENT in train sample",
     xlab = "DOCUMENT")
```

<br>
<br>

# 16. EXT_SOURCE_1

```{r}
# Summary statistics of EXT_SOURCE_1
summary(app_test3$EXT_SOURCE_1)
summary(app_train3$EXT_SOURCE_1)

sum(is.na(app_test3$EXT_SOURCE_1))
sum(is.na(app_train3$EXT_SOURCE_1))

# Create a subset of the data without missing values
no_missing_train <- app_train3$EXT_SOURCE_1[!is.na(app_train3$EXT_SOURCE_1)]
no_missing_test <- app_test3$EXT_SOURCE_1[!is.na(app_test3$EXT_SOURCE_1)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$EXT_SOURCE_1)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$EXT_SOURCE_1)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$EXT_SOURCE_1[is.na(app_train3$EXT_SOURCE_1)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$EXT_SOURCE_1[is.na(app_test3$EXT_SOURCE_1)] <- random_values_test

# Plot the distribution of EXT_SOURCE_1
hist(app_test3$EXT_SOURCE_1, 
     main = "Distribution of EXT_SOURCE_1 in test sample",
     xlab = "EXT_SOURCE_1")

hist(app_train3$EXT_SOURCE_1, 
     main = "Distribution of EXT_SOURCE_1 in train sample",
     xlab = "EXT_SOURCE_1")
```

<br>
<br>

# 17. EXT_SOURCE_2

```{r}
# Summary statistics of EXT_SOURCE_2
summary(app_test3$EXT_SOURCE_2)
summary(app_train3$EXT_SOURCE_2)

sum(is.na(app_test3$EXT_SOURCE_2))
sum(is.na(app_train3$EXT_SOURCE_2))

# Create a subset of the data without missing values
no_missing_train <- app_train3$EXT_SOURCE_2[!is.na(app_train3$EXT_SOURCE_2)]
no_missing_test <- app_test3$EXT_SOURCE_2[!is.na(app_test3$EXT_SOURCE_2)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$EXT_SOURCE_2)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$EXT_SOURCE_2)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$EXT_SOURCE_2[is.na(app_train3$EXT_SOURCE_2)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$EXT_SOURCE_2[is.na(app_test3$EXT_SOURCE_2)] <- random_values_test

# Plot the distribution of EXT_SOURCE_2
hist(app_test3$EXT_SOURCE_2, 
     main = "Distribution of EXT_SOURCE_2 in test sample",
     xlab = "EXT_SOURCE_2")

hist(app_train3$EXT_SOURCE_2, 
     main = "Distribution of EXT_SOURCE_2 in train sample",
     xlab = "EXT_SOURCE_2")
```

<br>
<br>

# 18. EXT_SOURCE_3

```{r}
# Summary statistics of EXT_SOURCE_3
summary(app_test3$EXT_SOURCE_3)
summary(app_train3$EXT_SOURCE_3)

sum(is.na(app_test3$EXT_SOURCE_3))
sum(is.na(app_train3$EXT_SOURCE_3))

# Create a subset of the data without missing values
no_missing_train <- app_train3$EXT_SOURCE_3[!is.na(app_train3$EXT_SOURCE_3)]
no_missing_test <- app_test3$EXT_SOURCE_3[!is.na(app_test3$EXT_SOURCE_3)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$EXT_SOURCE_3)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$EXT_SOURCE_3)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$EXT_SOURCE_3[is.na(app_train3$EXT_SOURCE_3)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$EXT_SOURCE_3[is.na(app_test3$EXT_SOURCE_3)] <- random_values_test

# Plot the distribution of EXT_SOURCE_3
hist(app_test3$EXT_SOURCE_3, 
     main = "Distribution of EXT_SOURCE_3 in test sample",
     xlab = "EXT_SOURCE_3")

hist(app_train3$EXT_SOURCE_3, 
     main = "Distribution of EXT_SOURCE_3 in train sample",
     xlab = "EXT_SOURCE_3")
```

<br>
<br>

# 19. AMT_REQ_CREDIT_BUREAU_QRT

```{r}
# Summary statistics of AMT_REQ_CREDIT_BUREAU_QRT
summary(app_test3$AMT_REQ_CREDIT_BUREAU_QRT)
summary(app_train3$AMT_REQ_CREDIT_BUREAU_QRT)

sum(is.na(app_test3$AMT_REQ_CREDIT_BUREAU_QRT))
sum(is.na(app_train3$AMT_REQ_CREDIT_BUREAU_QRT))

# Create a subset of the data without missing values
no_missing_train <- app_train3$AMT_REQ_CREDIT_BUREAU_QRT[!is.na(app_train3$AMT_REQ_CREDIT_BUREAU_QRT)]
no_missing_test <- app_test3$AMT_REQ_CREDIT_BUREAU_QRT[!is.na(app_test3$AMT_REQ_CREDIT_BUREAU_QRT)]

# Generate random values from the train subset distribution
random_values_train <- sample(no_missing_train, sum(is.na(app_train3$AMT_REQ_CREDIT_BUREAU_QRT)), replace = TRUE)
random_values_test <- sample(no_missing_test, sum(is.na(app_test3$AMT_REQ_CREDIT_BUREAU_QRT)), replace = TRUE)

# Replace missing values with random values in the train dataset
app_train3$AMT_REQ_CREDIT_BUREAU_QRT[is.na(app_train3$AMT_REQ_CREDIT_BUREAU_QRT)] <- random_values_train

# Replace missing values with random values in the test dataset
app_test3$AMT_REQ_CREDIT_BUREAU_QRT[is.na(app_test3$AMT_REQ_CREDIT_BUREAU_QRT)] <- random_values_test

# Plot the distribution of AMT_REQ_CREDIT_BUREAU_QRT
hist(app_test3$AMT_REQ_CREDIT_BUREAU_QRT, 
     main = "Distribution of AMT_REQ_CREDIT_BUREAU_QRT in test sample",
     xlab = "AMT_REQ_CREDIT_BUREAU_QRT")

hist(app_train3$AMT_REQ_CREDIT_BUREAU_QRT, 
     main = "Distribution of AMT_REQ_CREDIT_BUREAU_QRT in train sample",
     xlab = "AMT_REQ_CREDIT_BUREAU_QRT")
```
<br>
<br>

Change to binary variable, where if there were a request it is 1, if not - 0.

```{r}
app_test3$AMT_REQ_CREDIT_BUREAU_QRT <- ifelse(app_test3$AMT_REQ_CREDIT_BUREAU_QRT > 0, 1, 0)
app_train3$AMT_REQ_CREDIT_BUREAU_QRT <- ifelse(app_train3$AMT_REQ_CREDIT_BUREAU_QRT > 0, 1, 0)

hist(app_test3$AMT_REQ_CREDIT_BUREAU_QRT, 
     main = "Distribution of AMT_REQ_CREDIT_BUREAU_QRT in test sample",
     xlab = "AMT_REQ_CREDIT_BUREAU_QRT")

hist(app_train3$AMT_REQ_CREDIT_BUREAU_QRT, 
     main = "Distribution of AMT_REQ_CREDIT_BUREAU_QRT in train sample",
     xlab = "AMT_REQ_CREDIT_BUREAU_QRT")
```

```{r}
str(app_train3)

# Load the corrplot package
library(corrplot)

# Calculate correlation matrix
cor_matrix <- cor(app_train3)

# Create heatmap
heatmap(cor_matrix, 
        cmap = colorRampPalette(c("blue", "white", "red")), 
        main = "Correlation Matrix")
```


```{r}
app_train4<-app_train3
app_test4<-app_test3

#Karolina 
# file_path <- "C:\\Users\\karla\\OneDrive\\Pulpit\\Credit Risk\\Projekt\\app_train4.csv"

# Save the dataframe as a CSV file
# write_csv(app_train4, file_path)
# file_path <- "C:\\Users\\karla\\OneDrive\\Pulpit\\Credit Risk\\Projekt\\app_test4.csv"
# write_csv(app_test4, file_path)

save(app_train4, file = "dane/app_train4.RData")
save(app_test4, file = "dane/app_test4.RData")

```

